---
title: "Prediction Assignment"
author: "Rajesh Prabhakar"
output:
  html_document:
    df_print: paged
  keep_md: yes
---


# 1) Load & process training data

NA/empty columns & columns which contain data which doesn't come from accelerometers are removed.

```{r,echo = TRUE,message = FALSE}

library(dplyr)

training_data <- read.csv("pml-training.csv",na.strings = c("","NA"))

training_data <- training_data %>%
  select_if(~all(!is.na(.)))

training_data <- training_data[ ,-1:-7]

```

# 2) Split training data into training set & testing set

Cross validation will be done using random subsampling to create training & testing sets.

```{r,echo = TRUE,message = FALSE}

library(caret)

set.seed(123)

inTrain <- createDataPartition(y = training_data$classe,p = 0.75,list = FALSE)

training <- training_data[inTrain,]
testing <- training_data[-inTrain,]

```

# 3) Train model using training set

Model training will be done using random forests method on all accelerometer data. To reduce training time, parallel processing will be used & results will be cached.

```{r,echo = TRUE,message = FALSE,cache = TRUE}

library(parallel)
library(doParallel)

set.seed(123)

cluster <- makeCluster(detectCores())
registerDoParallel(cluster)
fitControl <- trainControl(allowParallel = TRUE)

modFit <- train(classe ~ . ,data = training,method = "rf",trcontrol = fitControl)

stopCluster(cluster)
registerDoSEQ()

```

# 4) Apply trained model to testing set

Out of sample error will be calculated using confusion matrix.

```{r,echo = TRUE,message = FALSE}

set.seed(123)

pred <- predict(modFit,testing)
testing$classe <- as.factor(testing$classe)
confusionMatrix(pred,testing$classe)

```

Accuracy is 0.995 & Kappa is 0.994 when trained model is applied to testing set, which indicates prediction model is very accurate.

# 5) Load & process testing data

NA/empty columns & columns which contain data which doesn't come from accelerometers are removed.

```{r,echo = TRUE,message = FALSE}

testing_data <- read.csv("pml-testing.csv",na.strings = c("","NA"))

testing_data <- testing_data %>%
  select_if(~all(!is.na(.)))

testing_data <- testing_data[ ,-1:-7]

```

# 6) Predict "classe" variable in testing data

```{r,echo = TRUE,message = FALSE}

pred <- predict(modFit,testing_data)
testing_data <- cbind(testing_data,pred)
print(testing_data[ ,c("problem_id","pred")])

```